???A quantitative framework for investigating the reliability of network construction

### 2017-09-26 manuscript/CI_interactions.tex
more small editsg

The basic information that we are collecting consists of interaction matrices of interaction partners. If our counts of whether species A is interacting with species B reflects both uncertainty with respect to whether two species interact when co-occurring at a given site, chance variation in whether the interaction is realized, and uncertainty with respect to whether it is detected when realized, then what can we do to pin down any of these different parts?

A simple consideration will illustrate the problem: assuming that we want to infer the probability of an interaction between two species, A and B. Now assume that in reality, interaction between A and B is plain impossible (i.e. the true (P(L|co-occurrence of A and B = 0). Given no added detection error in observing the incidence of the interaction, then a single observation of species co-occurrence will reveal (very) little regarding the probability of the interaction – the confidence limits for one observation essentially spans from 0 to 1 (see following section). For 5 observations, this confidence limit will still extends to 0.5. Only with thirty observations will it shrink to 0.1. Thus, adding (many) more observations is certainly useful in controlling uncertainty due to stochasticity inherent in the process itself.

The pattern of Fig. X (binomial CL fig from Dom) should be compared to the frequency of co-occurrence of taxa in the Salix-galler data set. Here, most(TO COMPUTE) pairs of species are never found co-occurring, and for species found co-occurring the number of co-occurrences was generally low (Fig from AC and KW). Among the species which did sometimes interact when co-occurring, the incidence of interaction was actually low(TO COMPUTE) (freq histogram Fig from AC and KW). Thus, even in the most extensive data set that we could think of, limited sampling is clearly a major source of uncertainty.

>>>>

The basic information that we are collecting consists of matrices of interaction partners. Our count of interactions between species A and species B reflects uncertainty with respect to whether two species interact when co-occurring at a given site, chance variation in whether the interaction is realized, and uncertainty with respect to whether it is detected when realized. In this case, what can we do to pin down any of these different types of variation?

A simple example will illustrate the problem: assume that we want to infer the probability of an interaction between two species, A and B. Now assume that in reality, interaction between A and B is completely impossible (i.e. the true (P(L|co-occurrence of A and B = 0). Given no added detection error in observing the incidence of the interaction, then a single observation of species co-occurrence will reveal (very) little regarding the probability of the interaction – the confidence limits for one observation essentially spans from 0 to 1 (see following section). For 5 observations, this confidence limit will still extends to 0.5. Only with thirty observations will it shrink to 0.1. Thus, adding (many) more observations is certainly useful in controlling uncertainty due to stochasticity inherent in the process itself.

\begin{figure}
  \caption{This figure illustrates the upper bound (solid black line) of a 95\% confidence interval for the interaction probability $\lamba$ when there is no observation of an interaction ($k = 0$) and $n$ observations of the two species co-occurring at a given location. The confidence interval is based on the Clopper-Pearson derivation from the beta distribution. A threshold interaction probability of 0.1 is indicated by the dotted red line.}
  \label{upper_limits}
  \includegraphics*[width=.8\textwidth]{figures/upper_limit_DG.eps}
\end{figure}

The pattern of Fig.~\ref{upper_limits} should be compared to the frequency of co-occurrence of taxa in the Salix-galler data set. Here, most(TO COMPUTE) pairs of species are never found co-occurring, and for species found co-occurring the number of co-occurrences was generally low (Fig from AC and KW). Among the species which did sometimes interact when co-occurring, the incidence of interaction was actually low(TO COMPUTE) (freq histogram Fig from AC and KW). Thus, even in the most extensive data set that we could think of, limited sampling is clearly a major source of uncertainty.
